Compact Particle-Mesh $N$-body Algorithm

### 2016-09-12 ms_caf/ms_caf.tex
haoran

\subsection{Spacial decomposition}
{\tt CUBE} uses cubic decomposition structures. The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes, and each of these are assigned to a coarray {\it image}\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran. We use this terminology in this paper.}. Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}. Each tile is surrounded by a {\it buffer} region which is usually ${\tt ncb}$ coarse cells thick. The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tiles buffer region to its center, {\it physical} region. Thus, a integer-based coarse mesh density is defined as
\begin{equation*}
{\tt integer(1)\ rho\_c(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
\end{equation*}
where ${\tt nex}={\tt nt}+2{\tt ncb}$ covers the buffer region on both sides, {\tt nnt} is the tile dimemsions, and {\tt nn} is the image {\it codimensions}\footnote{Coarray Fortran concept. Codimensions can do communications between images.}. Particles' phase coordinates {\tt x} and {\tt v} are copied between tiles and images, when necessary.
\begin{figure}[h]
\begin{verbatim}
    program CUBE
      call initialize
      call particle_initialize
      sync all
      call buffer_density
      call buffer_x
      call buffer_v
      do
        call timestep
        call update_x
        call buffer_density
        call buffer_x
        call PM ! update_v
        call buffer_v
        if(checkpoint_step) then
           call checkpoint
           if (final_step) exit
        endif
      enddo
      call finalize
    endprogram
\end{verbatim}
\caption{Overall structure of {\tt CUBE}.}
\label{fig.code}
\end{figure}

@@

Figure \ref{fig.code} shows the overall structure of the code. Subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation. Force kernels are also created or read in here. In {\tt particle\_initialize}, for each image, we read in initial positions and velocities in the compressed format. When this step is done by all images (synchronized by ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), the density fields {\tt rho\_c} are buffered between tiles (communications between images are needed) by subroutine {\tt buffer\_density}. Then, particles' phase space coordinates {\tt x} and {\tt v} are also buffered, by {\tt buffer\_x} and {\tt buffer\_v} respectively. In these steps, {\tt x} and {\tt v}'s ordering is changed according to the updated {\tt rho\_c}.

@@

\subsection{Memory layout}

\section{Results}

\section{Discussion and Conclusion}

\appendix
\section{Velocity distribution function}

%\begin{figure} \centering
%  \includegraphics[width=1.0\linewidth]{delta_mesh.pdf}
%  \caption{Caption goes here.}
%  \label{fig.1}
%\end{figure}

>>>>

\begin{figure}[t]
\includegraphics[width=1.0\linewidth]{fig_tile.pdf}
\caption{Spacial decomposition.}
\label{fig.tile}
\end{figure}

\subsection{Spacial decomposition}
{\tt CUBE} uses cubic decomposition structures. The global simulation volume is decomposed into ${\tt nn}^3$ cubic sub-volumes, and each of these are assigned to a coarray {\it image}\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran. We use this terminology in this paper.}. Inside of an image, the sub-volume is further decomposed into ${\tt nnt}^3$ cubic {\it tiles}. Each tile is surrounded by a {\it buffer} region which is usually ${\tt ncb}$ coarse cells thick. The buffer is designed for two reasons: (1) computing the fine mesh force, whose cut-off ${\tt nforce\_cutoff}\leq{\tt ncb}$, and (2) collecting all possible particles travelling from a tiles buffer region to its center, {\it physical} region. Thus, a integer-based coarse mesh density is defined as
\begin{equation*}
{\tt integer(1)\ rho\_c(nex,nex,nex,nnt,nnt,nnt)[nn,nn,*]}\\
\end{equation*}
where ${\tt nex}={\tt nt}+2{\tt ncb}$ covers the buffer region on both sides, {\tt nnt} is the tile dimemsions, and {\tt nn} is the image {\it codimensions}\footnote{Coarray Fortran concept. Codimensions can do communications between images.}. Particles' phase coordinates {\tt x} and {\tt v} are copied between tiles and images, when necessary. Figure \ref{fig.tile} shows the spacial decomposition in a 2-dimensional analogy, with ${\tt nn=2}$ and ${\tt nnt=2}$.

\begin{figure}[t]
\begin{verbatim}

    program CUBE
      call initialize
      call particle_initialize
      sync all
      call buffer_density
      call buffer_x
      call buffer_v
      do
        call timestep
        call update_x
        call buffer_density
        call buffer_x
        call PM ! update_v
        call buffer_v
        if(checkpoint_step) then
           call checkpoint
           if (final_step) exit
        endif
      enddo
      call finalize
    endprogram
\end{verbatim}
\caption{Overall structure of {\tt CUBE}.}
\label{fig.code}
\end{figure}

@@

Figure \ref{fig.code} shows the overall structure of the code. Subroutine {\tt initialize} creates necessary FFT plans and read in configuration files telling the program at which redshifts we need to do checkpoints, halofinds, or stop the simulation. Force kernels are also created or read in here. In {\tt particle\_initialize}, for each image, we read in initial positions and velocities in the compressed format.  A initial condition generator is also written in the compressed format (see Appendix B). When this step is done by all images (synchronized by ``{\tt sync all}'', which is equivalent to a {\tt mpi\_barrier}), the density fields {\tt rho\_c} are buffered between tiles (communications between images are needed) by subroutine {\tt buffer\_density}. Then, particles' phase space coordinates {\tt x} and {\tt v} are also buffered, by {\tt buffer\_x} and {\tt buffer\_v} respectively. In these steps, {\tt x} and {\tt v}'s ordering is changed according to the updated {\tt rho\_c}.

@@

\subsection{Memory layout}
Here we summarize the memory usage in unit of bpp (byte per particle).

\section{Results}
We use the same initial conditions, and use {\tt CUBEP3M} and {\tt CUBE} to simulate the large scale structure formation separately and compare their results.

Results of {\tt izipx=1}

Results of {\tt izipx=2}

\begin{figure}[h]
\begin{verbatim}

\end{verbatim}

\caption{Power spectrum differential plot}
\label{fig.power}
\end{figure}

\section{Discussion and Conclusion}
Cosmological simulation, PM method.

PID.

Summarize memory.

Phi, GPU.

\clearpage

\appendix
\section{A. Velocity distribution function}

\section{B. Initial condition generator}
